Linting Linearity in Core/System FC

### 2023-01-10 report.tex
Nova thesis lightweight template
Improve thesis theme
Extend introduction on Haskell's linear types

\documentclass[a4paper, draft]{report}

@@

\begin{document}
\maketitle
\tableofcontents

@@

Linear type systems\cite{} add expressiveness to existing type systems by
stating that linear resources must be used exactly once (linearly). In
programming languages with a linear type system, not using certain resources or
using them twice are type errors. Linear types can, thus, be used to, e.g.,
guarantee that file handles, socket descriptors, and allocated memory are freed
exactly once (leaks and double-frees become type errors)\cite{}, implement safe
high-performance language interoperability\cite{}, guarantee that quantum
entangled states are not duplicated~\cite{}, handle mutable state\cite{},
something on session types guarantees\cite{}, and much more.

@@

Despite their great promise and research literature, few programming languages
have linear types --
% , with Rust\cite{} (with its ownership types), Idris\{}, and recently
% Haskell\cite{} being the most popular.
it is hard to balance linear and non-linear types in a pratical language. Linear
Haskell\cite{linearhaskell} tackles this difficulty and describes a type system
to retrofit linear types into Haskell, a purely-functional lazy language, while
retaining backwards-compatibility and code reuse across linear and non-linear
users of libraries.
% , and, e.g., avoid the required boilerplate threading of linear resources that we will get to know ahead.

@@

>>>>

\documentclass[]{lwnovathesis}

@@

\begin{document}

\frontmatter

\maketitle
\tableofcontents

\mainmatter

@@

Linear type systems\cite{} add expressiveness to existing type systems by
stating that linear resources must be used \emph{exactly once}. In programming
languages with a linear type system, not using certain resources or using them
twice are type errors. Linear types can thus be used to, for example, guarantee
that file handles, socket descriptors, and allocated memory are freed exactly
once (leaks and double-frees become type errors)\cite{}, implement safe
high-performance language interoperability\cite{}, guarantee that quantum
entangled states are not duplicated~\cite{}, handle mutable state safely\cite{},
something on session types guarantees\cite{}, and much more.

@@

Despite their great promise and research literature, it is hard to balance
linear and non-linear types in practice and, consequently, few programming
languages have linear type systems. Among them are Idris2\cite{}, a linearly and
dependently typed language based on Quantitative Type Theory, Rust\cite{}, a
language whose ownership types build on linear ones to guarantee memory safety
without garbage collection or reference counting, and, more recently,
Haskell\cite{}, a \emph{purely functional} and \emph{lazy} language.

Linearity in Haskell stands out from linearity in Rust and Idris2 for two
distinct reasons:

\begin{itemize}
    \item Linear types were only added to the language roughly \emph{31 years
        after} Haskell's inception. It is an especially difficult endeavour to
        add linear types to a well-established language with a large library
        ecosystem and many active users, rather than to develop the language
        from the ground up with linear types in mind, and the successful
        approach as implemented in GHC 9.0, the leading Haskell compiler, was
        based on Linear Haskell\cite{linearhaskell}, where a linear type system
        designed with retaining backwards-compatibility and allowing code reuse
        across linear and non-linear users of the same libraries in mind was
        described.
    \item 
\end{itemize}

% Besides Haskell's supporting linear
% types according to the said paper, Idris2\cite{} supports linear types in a
% dependently typed setting, Clean\cite{} has uniqueness types which are closely
% related to linear types, and Rust\cite{} has ownership types which build from
% linear types. 
% released in the Glasgow Haskell Compiler (GHC) version 9.0.
% , and, e.g., avoid the required boilerplate threading of linear resources that
% we will get to know ahead.

% Linear types were introduced in GHC 9.0

@@

% TODO:Use glossary
